\documentclass[12pt]{report}
\usepackage{graphicx}
\graphicspath{{Images/}}
\usepackage{textcomp} %for the copywrite symbol
\linespread{1.6} %sets lines to 1.5 spacing.  1.6 would be double spaced
\usepackage[margin=1.0in]{geometry}
\usepackage[semicolon,round,sort&compress,sectionbib,numbers]{natbib}  
\usepackage{chapterbib}  
\usepackage{amsmath}
\usepackage{subcaption}
\usepackage{appendix} %for appendices
\usepackage{float}



\usepackage{hyperref} %make stuff clickable
\hypersetup{
	colorlinks,
	citecolor=black,
	filecolor=black,
	linkcolor=black,
	urlcolor=black
}


\usepackage{physics}  %for bras and kets!
\newcommand{\angstrom}{\mbox{\normalfont\AA }} %makes \angstrom do its thing!
 
\begin{document}





\input{Additional_files/title_page.tex}


\renewcommand{\thepage}{\roman{page}}% Roman numerals for page counter
 
\addcontentsline{toc}{section}{Table of Contents}
\tableofcontents





\chapter*{Abstract}


\addcontentsline{toc}{section}{Abstract}

Electron energy loss spectroscopy (EELS) is an electron microscopy technique ideal for analyzing lightweight elements. The technique relies on using reference spectra to fingerprint and analyze results. These references can be obtained either through experiments  on standards or through simulation.  Simulations provide the ability to produce spectra for previously unstudied or sensitive materials, but rely on numerous approximations. Central to the simulation of EELS are the approximations treating core holes and shielding.  In this work, a new method of calculating the magnitude of core hole screening in the case of lithium materials is developed and implemented in density functional theory calculations. The method is used to verify the validity of performing electron energy loss spectroscopy at 30 keV to reduce beam damage.  EELS fine structure are calculated for metallic lithium, $\mathrm{Li_2O}$, and LiF and marked improvements in agreement between calculation and experiment are observed. The technique uses linear response theory to relate the electron density to the core hole shielding contribution.  This contribution is then implemented via a non integer core hole in final state rule density functional theory calculations.  The improvements enabled by the technique open the possibility of conducting increasingly quantitative EELS analysis.
\\

\chapter*{Abr\'eg\'e}
\addcontentsline{toc}{section}{Abr\'eg\'e}
\input{Additional_files/french_abstract.tex}



\chapter*{Acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}
Many people need to be acknowledged for helping make this work possible.  Foremost are my supervisors,  Professor Raynald Gauvin and  Professor George P. Demopoulos for all their support, guidance, and valuable discussions.  Equally essential to this work are Fr\'ed\'eric Voisard and Nicolas Brodusch who provided all the experimental results needed to verify the theoretical ones presented here. I would also like to thank Motoki Nakamura  for taking the time to read and help edit my thesis and Fr\'ed\'eric Voisard for assistance in writing the French abstract. Acknowledgement is also due to the agencies who helped fund this work: NSERC and IREQ.  \\

Support also came from a multitude of places in ways that were not always appreciated in the moment.  To this end, I would like to thank everyone who helped instill me with love and excitement for physics and science.  Whether it was in a classroom, a pub, or wandering the streets, those uttering the phrase ``Hey, that's an interesting problem," inspired me into making this work happen. Finally, I would like to thank my parents for knowing that this was my path well before I did, and supporting me every step of the way. 




\chapter*{Contribution of Authors}
\addcontentsline{toc}{section}{Contribution of Authors}
\input{Additional_files/contributions.tex}










\chapter{Introduction}
\renewcommand{\thepage}{\arabic{page}}% Arabic numerals for page counter
\setcounter{page}{1}% sets page number

\input{Chapters/introduction.tex}


\chapter{Literature Review}\label{literature_review}
\input{Chapters/EELS.tex}
\section{Density Functional Theory (DFT)}\label{dft_section}
\input{Chapters/DFT.tex}
\section{EELS Calculations}\label{ELNES_section}
\input{Chapters/EELScalculations.tex}
\chapter{Methods}\label{methods}
\input{Chapters/method.tex}
\chapter{Results and Discussion}\label{results}
\input{Chapters/Results.tex}
\chapter{Conclusion}\label{conclusion}
\input{Chapters/Conclusion.tex}

\begin{appendices}

\chapter{Calculation Details} \label{calc_details}

\input{Additional_files/calculation_details.tex}




\end{appendices}


\newpage
\bibliographystyle{unsrt}


\bibliography{EELS,DFT,EELScalculation,introduction,methods}


\end{document}